---
title: "Portfolio 3"
author: "Cassandra Richardson"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

drugs <- read.csv(here::here("C:/Users/cassr/GitHub/project3/data/drug-use-by-age.csv"))

library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggrepel)
library(ggpubr)
```


# **Cassandra Richardson's Portfolio Piece #3 - Data Analysis Report on Common Drug Use by Age**


# Step 1: Finding a dataset.

### The dataset I will be using for this project is "drug-use-by-age" from the FiveThirtyEight datasets. This data was sourced by the National Survey on Drug Use and Health from the Substance Abuse and Mental Health Data Archive. The data can be found here: https://github.com/fivethirtyeight/data/tree/master/drug-use-by-age 

### Dataset description: This dataset covers usage of 13 different drugs across 17 age groups. For each drug, there is information about the "use," which is the percentage of those in an age group that have used that drug in the last 12 months, and "frequency," which is the median number of times a user in that age group used that drug in the past 12 months. 


# Step 2: Develop a research plan.

### For this data analysis report, I want to examine the usage of three drugs: alcohol, marijuana, and hallucinogens. My rationale for choosing these drugs is that I expect the perceptions of each drug to vary across age. Alcohol is the most common drug in the US, and views on marijuana and hallucinogens have been changing in recent years as their therapeutic benefits are being recognized, with laws reflecting this through their legalization or decriminalization. 

### With these three variables, I want to examine trends in both use and frequency across age groups. 

### Hypotheses: 
1. Alcohol use will not significantly differ with age, but frequency will significantly differ by age. 
2. Marijuana use will be significantly different by age, but frequency will not significantly differ by age. 
3. Hallucinogen use will be significantly different by age, but frequency will not signficantly differ by age. 


# Step 3: Before we begin any analyses, let's examine what this dataset includes:

```{r exploring dogs}
#View(drugs)
summary(drugs)
```

### The "view()" function opens the full dataset file in a new tab, so we can visually examine what the data looks like. 

### The "summary()" function gives us the min, max, 1st quartile, mean, and 3rd quartile for all quantitative variables. However, it looks like some of our variables are character variables when they should be numeric - we will need to change this in our data cleaning. 


# Step 4: Let's explore some more characteristics about the data:

```{r data characteristics}
glimpse(drugs)
head(drugs)
names(drugs)
missing <- !complete.cases(drugs)
drugs[missing]
```

### - glimpse() shows us an overview of the whole dataset. We see that there are 17 rows and 28 columns.
### - head() gives us the first 6 rows of data
### - names() tells us all of the variable names
### - missing <- !complete.cases() creates an object/tibble called "missing" with any missing observations across the entire dataset
### - drugs[missing] shows us that the "missing" tibble has 0 columns. This means that there was no missing data in this dataset, which makes it easier to work with! However, when I looked at the data through the "View()" function, I noticed there is at least one breed with values of "-" which will need to be removed during data cleaning.



# Step 5: Before we begin our analyses, let's **clean the data**! 

### First, I want to remove the values of "not found":

```{r removing}
drugs_clean <- na_if(drugs, "-")
#View(drugs_clean)
```

### Perfect! Looking at the new datset, we see that all values of "-" were replaced with NA. 

### Although we won't be using all drugs in our analyses, let's still fix them for the sake of cleaning the data and making it easier for future analyses. As such, let's change the character variables to numeric variables: 

```{r numeric}
sapply(drugs_clean, class)

drugs_clean$cocaine.frequency <- as.numeric(drugs_clean$cocaine.frequency) 
drugs_clean$crack.frequency <- as.numeric(drugs_clean$crack.frequency) 
drugs_clean$heroin.frequency <- as.numeric(drugs_clean$heroin.frequency) 
drugs_clean$inhalant.frequency <- as.numeric(drugs_clean$inhalant.frequency)
drugs_clean$oxycontin.frequency <- as.numeric(drugs_clean$oxycontin.frequency) 
drugs_clean$meth.frequency <- as.numeric(drugs_clean$meth.frequency) 

sapply(drugs_clean, class)
```

### For the sake of having a smaller, more targeted dataset, let's subset "drugs_clean" to only include our variables of interest:

```{r cleaning more}
drugs_cleaner <- subset(drugs_clean, select=c(age, n, alcohol.use, alcohol.frequency, marijuana.use, marijuana.frequency, hallucinogen.use, hallucinogen.frequency))
```

# Data cleaning is complete! To clean the data, I did the following steps:
1. Removed missing values
2. Converted character variables to numeric variables
3. Created a subset of the variables we will be using




# Step 6: Now that we have clean data, I want to get mean values for our variables of interest, regardless of age: 
```{r descriptives}
remove_age <- subset(drugs_cleaner, select = -c(age, n))

mean_drugs <- sapply(remove_age, mean, na.rm = T)

print(mean_drugs)
```


# Step 7: Now let's visualize our data using scatterplots.

```{r alcohol scatterplot}
ggplot(drugs_cleaner) +
  aes(x = alcohol.use,
      y = alcohol.frequency,
      fill = age, 
      color = age) +
  geom_jitter(alpha = .6) +
  labs(
    title = "Alcohol Use by Age in Past 12 Months",
    x = "Percentage in Age Group that Used Alcohol",
    y = "Median # Times Used Alcohol") +
  geom_text_repel(aes(label = age),
                  box.padding = .5,
                  direction = "y",
                  size = 2.5,
                  max.overlaps = Inf)
```

```{r marijuana scatterplot}
ggplot(drugs_cleaner) +
  aes(x = marijuana.use,
      y = marijuana.frequency,
      fill = age, 
      color = age) +
  geom_jitter(alpha = .6) +
  labs(
    title = "Marijuana Use by Age in Past 12 Months",
    x = "Percentage in Age Group that Used Marijuana",
    y = "Median # Times Used Marijuana") +
  geom_text_repel(aes(label = age),
                  box.padding = .5,
                  direction = "y",
                  size = 2.5,
                  max.overlaps = Inf)
```

```{r hallucinogen scatterplot}
ggplot(drugs_cleaner) +
  aes(x = hallucinogen.use,
      y = hallucinogen.frequency,
      fill = age, 
      color = age) +
  geom_jitter(alpha = .6) +
  labs(
    title = "Hallucinogen Use by Age in Past 12 Months",
    x = "Percentage in Age Group that Used Hallucinogens",
    y = "Median # Times Used Hallucinogens") +
  geom_text_repel(aes(label = age),
                  box.padding = .5,
                  direction = "y",
                  size = 2.5,
                  max.overlaps = Inf)
```


# Step 8: Let's perform the Chi-square test of independence to see if there are the relationships between drug use and age, and drug frequency and age, per my hypotheses:

```{r alcohol use and age correlation}
alc_use_chisq <- chisq.test(drugs_cleaner$alcohol.use, drugs_cleaner$age)

alc_use_chisq
```

```{r alcohol frequency and age correlation}
alc_freq_chisq <- chisq.test(drugs_cleaner$alcohol.frequency, drugs_cleaner$age)

alc_freq_chisq
```

```{r marijuana use and age correlation}
marij_use_chisq <- chisq.test(drugs_cleaner$marijuana.use, drugs_cleaner$age)

marij_use_chisq
```

```{r marijuana frequency and age correlation}
marij_freq_chisq <- chisq.test(drugs_cleaner$marijuana.frequency, drugs_cleaner$age)

marij_freq_chisq
```

```{r hallucinogen use and age correlation}
hallu_use_chisq <- chisq.test(drugs_cleaner$hallucinogen.use, drugs_cleaner$age)

hallu_use_chisq
```

```{r hallucinogen frequency and age correlation}
hallu_freq_chisq <- chisq.test(drugs_cleaner$hallucinogen.frequency, drugs_cleaner$age)

hallu_freq_chisq
```

### Unfortunately, none of our chi-squares tests were signficant (*p* > .05). This may be due to the way the age groups were coded, or an error in the test I used.

# Step 9: To explore if the lack of significance was due to the coding of age groups, or if the relationship only exists among those 12-21 years old, let's filter the data to that age range and re-run our tests. 

```{r subset to only those under 21}
drugs_young <- subset(drugs_cleaner, age <= "21")
```

```{r alcohol use and young age correlation}
alc_use_chisq_y <- chisq.test(drugs_young$alcohol.use, drugs_young$age)

alc_use_chisq_y
```

```{r alcohol frequency and young age correlation}
alc_freq_chisq_y <- chisq.test(drugs_young$alcohol.frequency, drugs_young$age)

alc_freq_chisq_y
```

```{r marijuana use and young age correlation}
marij_use_chisq_y <- chisq.test(drugs_young$marijuana.use, drugs_young$age)

marij_use_chisq_y
```

```{r marijuana frequency and young age correlation}
marij_freq_chisq_y <- chisq.test(drugs_young$marijuana.frequency, drugs_young$age)

marij_freq_chisq_y
```

```{r hallucinogen use and young age correlation}
hallu_use_chisq_y <- chisq.test(drugs_young$hallucinogen.use, drugs_young$age)

hallu_use_chisq_y
```

```{r hallucinogen frequency and young age correlation}
hallu_freq_chisq_y <- chisq.test(drugs_young$hallucinogen.frequency, drugs_young$age)

hallu_freq_chisq_y
```

### Unfortunately, none of these tests were signficant either (*p* > .05). If I used the correct test, this means that there is not a signficant correlation betwwen age and alcohol, marijuana, or hallucinogen use or frequency. 

#### *Note for Dr. Wiernik: For this correlation test, I chose to use the Chi-square test because I wanted to compare age (categorical variable) to drug use or frequency (numeric variable). I tried running a one-way ANOVA as well, but those tests did not produce any p values, which makes me believe that that test would not work either. I have no idea what other analytical tests I could run, and this assignment is making me realize I need to take more statistics classes.*


# Step 10: Back to the drawing board. I'm curious if use (percentage of people who have used a substance in the past 12 months) and frequency (median number of times that substance was used in the past 12 months) are correlated, to provide insight into addiction or dependance on a substance. 

### New hypotheses:
1. Alcohol use and frequency will be correlated.
2. Marijuana use and frequency will be correlated. 
3. Hallucinogen use and frequency will NOT be correlated. 

```{r alcohol correlations}
cor.test(drugs_cleaner$alcohol.use,
         drugs_cleaner$alcohol.frequency,
         method = "pearson")
```

```{r marijuana correlations}
cor.test(drugs_cleaner$marijuana.use,
         drugs_cleaner$marijuana.frequency,
         method = "pearson")
```

```{r hallucinogen correlations}
cor.test(drugs_cleaner$hallucinogen.use,
         drugs_cleaner$hallucinogen.frequency,
         method = "pearson")
```

### Finally. We have significant results. Alcohol use and frequency are signficantly correlated (r = .90, p < .001), marijuana use and frequency are signficantly correlated (r = .62, p < .01), and hallucinogen use and frequency are NOT signficantly correlated, although it is trending towards a negative correlation (r = -.43, p = .08). 


# **Interpretation of the Findings**

### Based on our analyses, there does not seem to be a significant difference in either alcohol, marijuana, or hallucinogen use or frequency by age. This suggests that these common drugs may be used in similar amounts across age groups (*although our plots don't necessarily suggest this?*). However, there is a signficant correlation between the percentage of the population that use alcohol or marijuana within the last 12 months and the median number of times that these substances are used. To me, this suggests that people are not just trying these substances once; instead, they are using them frequently. On the other hand, hallucinogen use and frequency was not signficant, although it was trending towards a negative correlation. This means that although people may try hallucinogens, they don't use them too often. 

